\section{Implementation}
The reference code is implemented in C++17 with OpenMP and FFTW~3.3.
Periodic boundaries are enforced by wrapping particle positions and grid indices modulo $(N_x,N_y)$.
The unit domain is $[0,1]^2$ with cell volume $\Delta x\,\Delta y$.

\subsection{Software Architecture}
Three executables are provided:
\begin{itemize}
    \item \texttt{pic\_sim}: full electrostatic PIC timestep loop for interactive runs.
    \item \texttt{pic\_benchmark}: sweeps the deposition parameter matrix and writes CSV results.
    \item \texttt{pic\_validation}: runs charge, spectral-noise, and Langmuir-wave diagnostics.
\end{itemize}
Deposition kernels share a \texttt{DepositionConfig} structure selectable at runtime by scheme, particle layout (AoS/SoA), sorting flag, thread count, and backend (CPU, GPU atomics, or GPU privatized).
Each scheme exposes paired AoS and SoA entry points plus OpenMP-parallel variants for $T>1$.

\subsection{Timestep Cycle}
The explicit leapfrog cycle is scheme-dependent during deposition.
For NGP, CIC, and TSC:
\begin{enumerate}
    \item Particle push: $\mathbf{x}_p \mathrel{+}= \mathbf{v}_p\,\Delta t$ with periodic wrap.
    \item Charge deposition at the new particle position.
    \item Poisson solve: FFT-based spectral solve computes $\phi$ and $\mathbf{E}$.
    \item Field gather: CIC interpolation of $\mathbf{E}$, then $\mathbf{v}_p \mathrel{+}= -(q_p/m_p)\mathbf{E}\,\Delta t$.
\end{enumerate}
For Esirkepov, deposition must use the pre-push position and the displacement $\mathbf{v}_p\,\Delta t$:
\begin{enumerate}
    \item Charge deposition along $[\mathbf{x}_p,\mathbf{x}_p+\mathbf{v}_p\Delta t]$.
    \item Particle push.
    \item Poisson solve and field gather as above.
\end{enumerate}
This ordering preserves trajectory-based charge conservation.

\subsection{Diagnostics and Reproducibility}
After deposition, we compute charge conservation error
\begin{equation}
    \epsilon_Q = \frac{\left|\sum_{i,j}\rho_{i,j}\,\Delta x\,\Delta y - \sum_p q_p\right|}{\sum_p |q_p|},
\end{equation}
spectral noise as the fraction of FFT power above one-third of the Nyquist wavenumber in each dimension, and total energy $W = W_{\mathrm{kin}} + W_{\mathrm{field}}$.
Benchmark output is written as CSV under \texttt{results/}; \texttt{scripts/plot\_results.py} converts these files to the figures in this paper.

\subsection{Data and Code Availability}
The complete source code, build scripts, and archived benchmark CSVs are available at \url{https://github.com/ehocine/pic-deposition}.
A citable release with DOI will be deposited on Zenodo (\texttt{TODO-DOI}); see \texttt{CITATION.cff} in the repository root.
Figures and summary tables regenerate from the archived data in \texttt{data/benchmarks/} via \texttt{scripts/plot\_results.py}, including CPU results (\texttt{benchmark.csv}), GPU results (\texttt{benchmark\_gpu.csv}), and conservation study output (\texttt{conservation\_study.csv}).
GPU benchmarks are reproduced with \texttt{scripts/kaggle\_gpu\_benchmark.sh}; long-run conservation studies run via \texttt{./pic\_benchmark --conservation}.
