\section{Experimental Methodology}
\label{sec:methodology}

\subsection{Benchmark Matrix}
Microbenchmarks sweep a full factorial over:
\begin{itemize}
    \item Particle counts $N_p \in \{10^4, 10^5, 10^6\}$.
    \item Grid sizes $64^2$, $128^2$, and $256^2$ on the unit square.
    \item Schemes: NGP, CIC, TSC, Esirkepov.
    \item Layouts: AoS and SoA.
    \item Sorting: enabled and disabled.
\end{itemize}
Each case uses 5 timed repeats after 3 warmup depositions.
Particles are initialized from a uniform Maxwellian with fixed seed 42.
Esirkepov microbenchmarks use $\Delta t = 0.01$ to form trajectory weights.
CPU scaling cases fix $N_p=10^5$, grid $128^2$, CIC, SoA, no sorting, and thread counts $T \in \{1,2,4,8\}$ with privatized OpenMP buffers.
Energy-drift simulations run 500 steps for all schemes at $N_p=10^5$ on a $128^2$ grid with sorting enabled.

\subsection{Timing Protocol}
Timed deposition separates sort cost from kernel cost.
When sorting is enabled, one sort is timed, then the deposit kernel is averaged over \texttt{repeats} iterations on the sorted particle order.
Throughput is reported as $N_p / t_{\mathrm{deposit}}$ and effective bandwidth as $(B_{\mathrm{read}}+B_{\mathrm{write}})/t_{\mathrm{deposit}}$, where $B_{\mathrm{read}}$ counts particle record bytes and $B_{\mathrm{write}}$ counts stencil updates plus one grid clear.
Sort time is excluded from throughput and bandwidth so that locality effects can be compared independently.

\subsection{Physics Validation}
Langmuir-wave validation uses $N_p=200$ particles on a $64^2$ grid with $\Delta t=0.002$ and 16384 timesteps.
A sinusoidal velocity perturbation of amplitude 0.05 is imposed at mode number $m=1$.
After each deposition, a neutralizing background $\rho_{\mathrm{bg}} = 1/(L_x L_y)$ is added.
Field energy is recorded over the final three-quarters of the run and analyzed by discrete Fourier transform.
Pass criterion: measured frequency spread across all four schemes $\leq 5\%$.
We also report $\omega_{\mathrm{meas}}/\omega_{p,\mathrm{macro}}$ to document the systematic offset from the 1D cold-plasma estimate in this 2D diagnostic setup.

Energy-drift validation uses separate 500-step full-PIC simulations at $N_p=10^5$ on a $128^2$ grid.
A long-run conservation study executes 2000 full-PIC timesteps at $N_p=10^4$ on a $64^2$ grid for all four schemes, recording the maximum normalized charge error $|\epsilon_Q|$ and final relative total-energy drift.
This complements single-step microbenchmarks by exposing accumulated deposition--field coupling over thousands of leapfrog cycles.

\subsection{Roofline Methodology}
Roofline plots (Fig.~\ref{fig:roofline}) estimate arithmetic intensity as $(N_p \times \mathrm{FLOPs/particle}) / (B_{\mathrm{read}}+B_{\mathrm{write}})$ using the same byte model as effective-bandwidth reporting.
Stencil-dependent FLOP heuristics are NGP $\approx 10$, CIC $\approx 40$, and TSC $\approx 90$ floating-point operations per particle (approximate stencil arithmetic, not rigorous operation counts).
Achieved GFLOP/s is $(N_p \times \mathrm{FLOPs/particle}) / t_{\mathrm{deposit}}$.
Reference slopes correspond to memory-bandwidth plateaus observed on M1 Pro ($\sim$5--13\,GB/s) and Tesla T4 HBM ($\sim$300\,GB/s peak, lower achieved in scatter kernels).

\subsection{GPU Privatized Tile Strategy}
The \texttt{GPU\_Priv} backend launches one CUDA block per $16\times16$ spatial tile with a shared-memory accumulator ($2$\,KB per block on T4).
All threads in a block stride over the full particle list; stencil contributions landing in the tile are accumulated with shared-memory \texttt{atomicAdd}, then merged into global memory.
Each tile therefore revisits every particle, which is correct but $\mathcal{O}(N_{\mathrm{tiles}}\,N_p)$; reported privatized times reflect this cost after the charge-conservation fix.
CIC privatized tiles are benchmarked on all grid sizes; NGP and TSC remain atomics-only until tile helpers are verified.

\subsection{Hardware and Scope}
CPU benchmarks were built with \texttt{cmake -DCMAKE\_BUILD\_TYPE=Release} on Apple M1 Pro (Table~\ref{tab:hardware}).
GPU benchmarks used the same build flags with \texttt{BUILD\_CUDA=ON} and \texttt{CMAKE\_CUDA\_ARCHITECTURES=75} on dual Tesla T4 GPUs via Kaggle (Table~\ref{tab:gpu_hardware}); reported GPU times use a single device.
The GPU matrix compares CPU, \texttt{GPU\_Atomics}, and \texttt{GPU\_Priv} backends for NGP, CIC, and TSC on SoA layout without sorting, over the same particle counts and grid sizes as the CPU microbenchmarks.
All reported figures use deposit-kernel time unless explicitly labeled otherwise.

\begin{table}[t]
\centering
\small
\caption{CPU benchmark platform (Apple M1 Pro).}
\label{tab:hardware}
\begin{tabular}{ll}
\toprule
Component & Specification \\
\midrule
CPU & Apple M1 Pro \\
Cores & 8 (OpenMP threads: 1, 2, 4, 8) \\
OS & macOS 26.5.1 \\
Compiler & Apple clang 21.0.0, C++17, \texttt{-O3} \\
Libraries & FFTW 3.3.11, libomp 22.1.7 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[t]
\centering
\small
\caption{GPU benchmark platform (Kaggle, Tesla T4).}
\label{tab:gpu_hardware}
\begin{tabular}{ll}
\toprule
Component & Specification \\
\midrule
GPU & NVIDIA Tesla T4 (15\,GB) \\
CUDA & 12.8, \texttt{sm\_75} \\
Host CPU & Intel Xeon (Kaggle VM) \\
Compiler & GCC 11.4.0, C++17, \texttt{-O3} \\
Libraries & FFTW 3.3.8, OpenMP 4.5 \\
\bottomrule
\end{tabular}
\end{table}
